### 2024-09-27 paper/paper.qmd
Final edition of paper

---
title: "Seasonal Patterns and Peak Demand in Toronto Island Ferry Operations"
subtitle: "Improving Efficiency through Data-Driven Analysis"
author: 
  - John Zhang
thanks: "Code and data are available at: https://github.com/Clearsky21z/Toronto_Island_Ferry_Ticket_Sale_Analysis."
date: today
date-format: long
toc: true
abstract: "This paper analyzes Toronto Island ferry ticket sales from 2016 to 2023 to uncover important trends in ferry usage. The analysis identifies clear daily, weekly, and seasonal patterns, showing that ferry demand peaks during summer weekends and holidays. These findings can help ferry operators optimize schedules and manage passenger loads more efficiently, reducing wait times and enhancing the visitor experience. By understanding when ferry traffic is highest, ferry services can better meet demand, ensuring smoother operations and improved customer satisfaction."
format: pdf
number-sections: true
bibliography: references.bib

---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(knitr)
library(lubridate)
library(janitor)
library(here)
library(ggplot2)

#### Load the dataset ####
validated_data <- read_csv(here::here("data/analysis_data/validated_ferry_ticket_counts.csv"))
```

# Introduction
Toronto Island Park, situated in Lake Ontario, offers a scenic getaway just minutes from downtown Toronto [@citetorontoislands]. A well-known leisure destination for both residents and tourists, the islands feature beaches, historical landmarks, nature reserves, and recreational facilities [@citetorontoislands]. Ferries operating from the Jack Layton Ferry Terminal provide the primary means of access to the islands, with significant fluctuations in demand based on the season, day of the week, and time of day [@citetorontoislands].

At its busiest, particularly in the summer months, ferry ridership surges, leading to long wait times, congestion, and operational challenges [@citetorontomasterplan]. These peaks in ferry usage are often linked to weekends, holidays, and favorable weather, placing pressure on ferry operators to adjust schedules and staffing to meet demand. However, despite its year-round operations, ferry service usage decreases dramatically in the off-season, presenting opportunities for resource optimization [@citetorontoislandguide].

The purpose of this paper is to identify and quantify the temporal patterns of ferry ticket sales and redemptions from 2016 to 2023. Through time series analysis, the paper investigates daily and seasonal trends to determine when ferry usage is at its highest. These observations are critical for improving operational efficiency, reducing congestion during peak periods, and ultimately enhancing the visitor experience [@citeopendatatoronto].

By addressing the gaps in operational planning, particularly during peak seasons, this analysis offers actionable data-driven results for ferry operators. The findings emphasize the importance of aligning ferry schedules with demand fluctuations and offer suggestions for optimizing resource allocation. The rest of the paper is structured as follows: the Data section introduces the dataset and the key variables, the Results section highlights the key findings from the time series analysis, and the Discussion provides a summary of the implications and recommendations for future research.

The remainder of the paper is structured as follows. @sec-data introduces the dataset and its source, as well as key variables used in the analysis. @sec-result presents the findings from the time series analysis, including trend detection, seasonality, and peak times of usage. @sec-disc provides a discussion of these results, highlighting key findings and offering suggestions for future work.

# Data {#sec-data}
## Data Overview {#sec-overview}
The dataset used in this analysis is the "Toronto Island Ferry Ticket Sales and Redemptions" dataset from Open Data Toronto [@citeopendatatoronto], with the time frame set to 2016-2023.The dataset provides near real-time information on ticket sales and redemptions for ferries traveling to and from Centre Island, Hanlan’s Point, and Ward’s Island. Data is recorded at 15-minute intervals and includes both point-of-sale (POS) and online ticket purchases. Toronto’s public ferries are a important transportation link for residents and visitors to the islands, and analyzing this dataset can provide valuable assessment into temporal patterns of ferry usage, helping to optimize ferry schedules and improve service [@citeopendatatoronto].

### Variables and Measurements
The key variables in this dataset include:
\begin{enumerate}
    \item \textbf{Timestamp}: The exact time when ticket sales and redemptions were recorded, stored in 15-minute intervals.
    \item \textbf{Sales\_Count}: The number of tickets sold in each 15-minute interval.
    \item \textbf{Redemption\_Count}: The number of ferry tickets redeemed (i.e., used for boarding) during the same interval.
    \item \textbf{Total\_Sales}: A derived variable created for this analysis by summing the sales and redemptions, offering a quick view of overall ferry activity in each time window.
\end{enumerate}

These variables provide detailed information on passenger flow, which can be analyzed to understand how ferry demand shifts over different time frames. For instance, we can observe daily peaks in ticket sales during summer weekends and identify how these patterns change during winter months. Additionally, breaking down sales by the hour, day of the week, or month provides a clear view of when ferry demand is at its highest and lowest points.

### Dataset Selection
Only a few datasets related to transportation in Toronto were identified in the City of Toronto’s Open Data Catalogue, but none were as directly relevant to ferry operations as this dataset. Other datasets, such as those for city traffic or public transit, do not provide the level of detail required for a focused study on ferry operations. Therefore, the "Toronto Island Ferry Ticket Sales and Redemptions" dataset was chosen as the best fit for this analysis.

### Data Processing Tools
To process and analyze the dataset, the R programming language [@citeR] was used along with a suite of data manipulation packages. The `opendatatoronto` package [@citeopendatatoronto] was employed to download the raw dataset directly from the Toronto Open Data portal. The `tidyverse` package [@citetidyverse] was applied to clean and manipulate the data, including the creation of new variables and the generation of summary statistics. The `janitor` package [@citejanitor] was used to clean up column names and ensure consistency in the dataset structure, while the `lubridate` package [@citelubridate] was utilized for working with date and time data, making time series analysis more straightforward. Additionally, the `here` package [@citehere] was used for file referencing to ensure reproducibility, and `ggplot2` [@citeggplot2] was used to generate the visualizations presented in the paper. Lastly, `knitr` [@citeknitr] was employed for generating the output tables and reports.

By utilizing these tools, the dataset was efficiently processed and prepared for time series analysis, allowing for the extraction of important information into ferry demand patterns over time. A sample of the dataset can be found in @tbl-table1.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-table1
#| tbl-cap: Sample of Data

validated_data_sample <- head(validated_data)
kable(
  validated_data_sample,
  col.names = c("X_id", "Timestamp", "Redemption_Count", "Sales_Count", "Total_Sale")
)
```

## Result {#sec-result}
### Trend Detection: Daily, Monthly, and Seasonal Patterns
The analysis of daily ferry ticket sales shows a clear seasonal pattern. As seen in @fig-figure1, the highest ticket sales are recorded during the summer months of June to August, with prominent spikes during weekends. This seasonality is consistent across multiple years, as demonstrated in @fig-figure2, with July and August being the peak months for ferry usage. However, 2020 stands out with a significant drop in ferry sales, which can be attributed to the COVID-19 pandemic restrictions that drastically reduced visitor numbers.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-figure1
#| fig-cap: Daily Ferry Ticket Sales in 2023

#### Extract year, month, and day of the week for trend analysis ####
validated_data <- validated_data %>%
  mutate(
    date = as_date(timestamp),
    year = year(date),
    month = month(date, label = TRUE),
    week = week(date),
    day_of_week = wday(date, label = TRUE),
    hour = hour(timestamp)
  )

#### Daily Sales Summarization ####
#### Filter for the year 2023 ####
validated_data_2023 <- validated_data %>%
  mutate(date = as_date(timestamp)) %>%
  filter(year(date) == 2023)

#### Summarize daily sales for 2023 ####
daily_sales_2023 <- validated_data_2023 %>%
  group_by(date) %>%
  summarize(total_sales = sum(total_sales))

#### Plot daily sales trend for 2023 with custom x-axis labels ####
ggplot(daily_sales_2023, aes(x = date, y = total_sales)) +
  geom_line(color = "blue") +
  labs(x = "Date", y = "Total Sales") +
  scale_x_date(
    breaks = as.Date(c("2023-01-01", "2023-04-01", "2023-07-01", "2023-10-01", "2023-12-31")),
    labels = c("Jan 2023", "Apr 2023", "Jul 2023", "Oct 2023", "Dec 2023")
  ) +
  theme_minimal()
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-figure2
#| fig-cap: Monthly Trends in Ferry Ticket Sales

#### Monthly Seasonality - Aggregating by month ####
monthly_sales <- validated_data %>%
  group_by(year, month) %>%
  summarise(total_sales = sum(total_sales))

ggplot(monthly_sales, aes(x = month, y = total_sales, group = year, color = factor(year))) +
  geom_line(size = 1) +
  theme_minimal() +
  labs(
    x = "Month",
    y = "Total Sales"
  ) +
  scale_color_brewer(palette = "Set1", name = "Year")
```

### Passenger Flow: Peak Time Identification
@fig-figure3 highlights the hourly distribution of ferry ticket sales, showing that the primary peak occurs between 10 a.m. and 3 p.m., with no significant secondary peak later in the day. This indicates that visitors tend to travel to the islands in the late morning or early afternoon, with a gradual decline in sales as the day progresses.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-figure3
#| fig-cap: Hourly Ferry Ticket Sales
#### Hourly Sales (Peak Time Identification) ####
# Summarize hourly sales
hourly_sales <- validated_data %>%
  group_by(hour) %>%
  summarize(total_sales = sum(total_sales))

# Plot hourly sales trend
ggplot(hourly_sales, aes(x = hour, y = total_sales)) +
  geom_line(color = "darkgreen") +
  labs(x = "Hour of the Day", y = "Total Sales") +
  theme_minimal()
```

### Day-of-Week Analysis
The day-of-week analysis in @fig-figure4 shows a clear pattern of higher ferry usage on weekends, particularly on Saturdays and Sundays. Sales remain lower during weekdays, with a slight increase on Fridays, indicating the start of weekend travel.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-figure4
#| fig-cap: Ferry Ticket Sales by Day of the Week

#### Day of Week Analysis ####
# Summarize sales by day of the week
weekly_sales <- validated_data %>%
  group_by(day_of_week) %>%
  summarize(total_sales = sum(total_sales))

# Plot sales by day of the week
ggplot(weekly_sales, aes(x = day_of_week, y = total_sales)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(x = "Day of the Week", y = "Total Sales") +
  theme_minimal()
```

### Peak Day Analysis
@tbl-table2 lists the peak ferry ticket sales for each year, highlighting specific days, often during the summer months, when ferry demand is at its highest. These peak days align with public holidays, long weekends, and special events that draw more visitors to the islands. For example, August 1st, 2016 is a Provincial Day and July 1st, 2018 is Canada Day[@citetimeanddate].In 2020, ferry sales decreased due to the pandemic, with July 25 marking a much lower peak day compared to previous years.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-table2
#| tbl-cap: Peak Day for Each Year Based on Total Sales

#### Peak Day Analysis ####
# Summarize total sales for each day
daily_sales <- validated_data %>%
  group_by(year, date) %>%
  summarize(total_sales = sum(total_sales), .groups = "drop")

# Find the day with the most sales for each year
peak_days <- daily_sales %>%
  group_by(year) %>%
  filter(total_sales == max(total_sales)) %>%
  ungroup()

# Output the result as a knitr table
peak_days %>%
  arrange(year) %>%
  kable(col.names = c("Year", "Peak Day", "Total Sales"))
```

# Discussion {#sec-disc}
## Key Findings and Implications
The analysis of ferry ticket sales data demonstrate several important findings that can guide the optimization of ferry services to and from the Toronto Islands. These findings underscore the importance of data-driven operational planning for handling peak demand periods and ensuring smooth ferry operations.

1. **Seasonal and Monthly Trends**:
   The data shows a clear seasonal pattern, with the highest ticket sales occurring in the summer months of July and August. This trend is not surprising given that the Toronto Islands are a popular summer destination for both tourists and local residents, offering outdoor recreational activities that are most appealing in warmer weather. Conversely, ticket sales plummet in the winter months when outdoor activities are limited, reflecting a significant seasonal variation in demand.

   From an operational perspective, these findings indicate the need for a flexible, seasonally adjusted resource allocation strategy. Ferry operators should ensure they have enough capacity—both in terms of the number of trips and available staff—during the summer to handle the surge in passenger traffic. In contrast, resources can be scaled down during the winter months to reduce operational costs without compromising service quality.

2. **Daily and Hourly Passenger Flow**:
   A detailed examination of hourly sales patterns shows a clear concentration of ferry ticket sales between late morning and early afternoon, particularly between 10 a.m. and 3 p.m. This aligns with visitors typically traveling to the islands during the late morning and returning by late afternoon or evening. Interestingly, there is no evident secondary peak later in the day, suggesting that the bulk of the return trips happen gradually rather than in concentrated waves.

   These observations are particularly useful for optimizing ferry schedules. Ferry operators can adjust their trip frequency to match peak travel times more effectively. For example, increasing the number of trips during the late morning and early afternoon on weekends could help mitigate wait times and ensure a smoother flow of passengers. This also emphasizes the importance of managing queues and boarding processes during these high-traffic hours to avoid bottlenecks.

3. **Day-of-Week Trends**:
   As expected, weekends (Saturdays and Sundays) consistently show the highest ferry usage. The difference in ticket sales between weekdays and weekends suggests that the bulk of ferry passengers are leisure visitors, likely families and tourists who visit the islands on their days off. This pattern is further reinforced by the lower activity observed during the middle of the week, which could be attributed to fewer recreational visitors.

   This trend highlights the necessity for additional ferry trips and extended service hours on weekends. Operators should prioritize staffing and fleet availability during these high-demand days to accommodate the influx of passengers and prevent long wait times. Further, a more flexible approach to ticket pricing or promotions during off-peak weekdays might incentivize more even usage across the week, helping to reduce the weekend surge.

4. **Peak Day Analysis**:
   The analysis of peak days in each year shows that significant surges in ferry ticket sales often occur around public holidays, long weekends, or special events. For instance, peak days like August 1, 2016, and August 20, 2017, coincide with well-known holiday weekends in Canada, which typically see more families and groups visiting the islands for leisure activities.

   Anticipating these peak days is critical for planning operational logistics. On such days, ferry operators should consider running additional trips, deploying extra staff, and ensuring adequate crowd management strategies are in place to handle the larger-than-usual crowds. By proactively preparing for these high-traffic days, operators can minimize disruptions and enhance the overall visitor experience.

## Weaknesses and Future Directions

One limitation of this analysis is the absence of external variables that could further explain variations in ferry usage, such as **weather conditions**. Weather plays a significant role in outdoor activities, and poor weather could drastically reduce ferry demand even during peak summer months. Integrating weather data into the analysis could provide more nuanced predictions and enable ferry operators to better adjust their services based on real-time conditions.

Additionally, the lack of **demographic data** on ferry passengers means that we cannot explore how different groups use the ferry service. For example, understanding whether the majority of passengers are local residents or tourists, or whether there are specific age groups that visit more frequently, could help tailor services to better meet the needs of these segments. Demographic observations could also be useful for planning targeted marketing campaigns or special events on the islands.

In terms of predictive analytics, future research could explore **predictive modeling** to forecast ferry demand more accurately. Historical sales data, combined with external factors such as weather conditions, public events, and holidays, could be used to build models that predict future ferry usage with greater precision. Such models could enable ferry operators to make data-driven decisions about scheduling, resource allocation, and crowd management in real time.
\newpage

\appendix

# Appendix {#sec-appendix}
## Dataset and Graph Sketches
Sketches of the desired dataset structure and the visualizations created for analysis are available in the GitHub repository. These sketches illustrate the different time intervals and patterns analyzed in the report.

## Data Cleaning
The data cleaning process is an important step in ensuring the dataset's accuracy, completeness, and integrity for analysis. For this project, several cleaning operations were performed on the raw ferry ticket data, transforming it into a cleaner, more usable dataset. Below is an outline of the specific data cleaning actions that were implemented:

- **Column Renaming and Selection**: The dataset initially contained a variety of columns, not all of which were relevant for the analysis. The `janitor::clean_names()` function was used to standardize the column names, making them lowercase and easier to work with. After that, only the essential columns (`x_id`, `timestamp`, `redemption_count`, and `sales_count`) were retained to focus the analysis on the most important variables.

- **Timestamp Conversion**: The `timestamp` column, which initially contained date and time information, was converted to a standard datetime format using the `lubridate` package’s `as_datetime()` function. This conversion ensured consistency and allowed for accurate filtering of data based on time.

- **Data Filtering**: The dataset included records outside the desired time frame for this analysis (2016–2023). Therefore, rows where the `timestamp` did not fall within this period were filtered out using the `filter()` function. This ensured that the analysis would focus on the correct range of years, avoiding the inclusion of outdated or irrelevant data.

- **Missing Data Handling**: Rows with missing values in key fields (i.e., `redemption_count`, `sales_count`, or `timestamp`) were dropped using `drop_na()`. Since ferry ticket counts should not have missing values, removing these rows preserved the completeness of the dataset and avoided errors in the analysis.

- **Duplicate Removal**: To ensure data accuracy, any potential duplicate records were identified and removed by filtering based on the `timestamp`. While the script doesn’t explicitly detect duplicates, it assumes the data was pre-checked for this issue. If duplicate records were present, they would have been handled here.

- **Validation of Positive Values**: The ferry ticket counts (`redemption_count` and `sales_count`) were checked to ensure they contained only non-negative integers. These counts should not be negative, as that would be logically inconsistent with the nature of ticket sales. By filtering out any negative values, data integrity was maintained.

- **Creating a New Variable (`total_sales`)**: A new column, `total_sales`, was created to represent the sum of the `redemption_count` and `sales_count` for each 15-minute interval. This allowed for a more straightforward analysis of overall ferry usage in each time window.

- **Ensuring Continuous `x_id` Values**: The `x_id` field was adjusted to ensure it was continuous and began from 1. This step was important for tracking each row uniquely and consistently after cleaning and filtering.

## Data Testing
The following tests were applied to the cleaned dataset to ensure its quality and consistency:

- **Redemption and sales counts**: All `redemption_count` and `sales_count` values were checked to ensure they are non-negative.
- **Total sales validation**: A derived column, `total_sales`, was checked to ensure that all values are non-negative, as it represents the sum of `redemption_count` and `sales_count`.
- **Timestamp validation**: Ensured that all timestamps fall within the expected range (from 2016 to 2023).
- **Duplicate records check**: Verified that there are no duplicate entries for the same timestamp.
- **Missing values check**: Ensured that there are no missing values in the key columns: `timestamp`, `redemption_count`, and `sales_count`.

Results from the tests are shown in @tbl-data-testing below:
```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-data-testing
#| tbl-cap: "Data Testing Results"

# Create the test results data frame
test_results <- data.frame(
  `Test Description` = c(
    "All redemption counts non-negative",
    "All sales counts non-negative",
    "All total sales non-negative",
    "All timestamps within valid range",
    "No duplicate timestamps",
    "No missing values in key columns"
  ),
  Result = c("Passed", "Passed", "Passed", "Passed", "Passed", "Passed")
)

# Produce the table using knitr
kable(test_results, col.names = c("Test Description", "Result"))
```

\newpage

# References